---
title: "Data Handle"
author: "James Woods"
date: "1/14/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reading Data In

## Financial Data

This was downloaded from Oregon department of eduction.

```{r}
DistrictFinance <- read.csv("https://s3-us-west-2.amazonaws.com/woodsjeeschools/ActualExpendituresFundFunction30.CSV") 

library(stringr)
library(dplyr)

#Convert to starting of fiscal year.  Remember that the year may not match with the test scores.
DistrictFinance$YearBeginning <- as.factor(str_sub(as.character(DistrictFinance$SchlYr),-4) )

```


```{r}
summary(DistrictFinance)
```

## Test data read

```{r}
library(openxlsx)

# New combined files
Tests15 <- read.xlsx("https://www.oregon.gov/ode/educator-resources/assessment/TestResults2015/pagr_schools_tot_ela_math.xlsx")

Tests16  <- read.xlsx("https://www.oregon.gov/ode/educator-resources/assessment/TestResults2016/pagr_schools_tot_ela_math.xlsx")

#Older files by topic can't be opened by xlsx so saving in gdocs and reading via csv.  These were downloaded via https://www.oregon.gov/ode/educator-resources/assessment/Pages/Test-Results-0910-through-1314.aspx 

## Math

Math14 <- read.csv("https://docs.google.com/spreadsheets/d/1mKkhwPVO_J4p8RCBuKiJauYz8Kx7xdBx4Yq9xqlVjss/pub?gid=336032123&single=true&output=csv")

Math13 <- read.csv("https://docs.google.com/spreadsheets/d/107Lp_UBqcTnObunO1lIPhq0sBOrpfSryHdW1HYY6qm4/pub?gid=1450304332&single=true&output=csv")

Math12 <- read.csv("https://docs.google.com/spreadsheets/d/1imChJVXmq4iVeaeB2S8qCeM01GxVegOhwW6v4RBswTI/pub?gid=1330438594&single=true&output=csv")

Math11 <- read.csv("https://docs.google.com/spreadsheets/d/12Tqn4RORNI2w3KoiqnVpq_Trd3Zm1MM51I5YXDZFvfk/pub?gid=657209680&single=true&output=csv")

Math10 <- read.csv("https://docs.google.com/spreadsheets/d/1CXKQ_yzcvQfEcS9C3iBD3Hsvj4ZqJ6ISuFwgoghzEFI/pub?gid=2015290170&single=true&output=csv")

## Reading

Read14 <- read.csv("https://docs.google.com/spreadsheets/d/1yJrAFMCcrqB8C_hLRDr6pN1ZWU4KUPON1CMc3h3yeAM/pub?gid=1572187371&single=true&output=csv")
Read13 <- read.csv("https://docs.google.com/spreadsheets/d/1MBhoFCZg2nAV7cSQnK0QX4s8yF8Oeda-ceaqWpyA3DA/pub?gid=1052055153&single=true&output=csv")
Read12 <- read.csv("https://docs.google.com/spreadsheets/d/1msHh_WXdRAKTPeUP6OgvK9AzWlbSx5O7pMBntq83dSA/pub?gid=652537434&single=true&output=csv")
Read11 <- read.csv("https://docs.google.com/spreadsheets/d/1Oe_P_Pxi4eIWtEWZZNfftzGNCdqNSvckMSfMaxYDZA4/pub?gid=1449554276&single=true&output=csv")
Read10 <- read.csv("https://docs.google.com/spreadsheets/d/1DScEAY9FPLRv7QoQrIhjNYQdF6LY078ZvVUQmnB2jWA/pub?gid=541255152&single=true&output=csv")


## Science

Sci14 <- read.csv("https://docs.google.com/spreadsheets/d/1AhXN382JnOrDv3MMp6bBI640VpZTpf1uuI-gm-xep3M/pub?gid=747992313&single=true&output=csv")
Sci13 <- read.csv("https://docs.google.com/spreadsheets/d/1JeEBL4yMxqKbnTh43mi-L5HEyaXBO96iuzw9UNUi0dE/pub?gid=1005826807&single=true&output=csv")
Sci12 <- read.csv("https://docs.google.com/spreadsheets/d/1tBnFq7P9r65zGP9AFEvVBExUCrGJWh9FystJryk83TI/pub?gid=1064106863&single=true&output=csv")
Sci11 <- read.csv("https://docs.google.com/spreadsheets/d/1hAJjcuFxSbLyWZ_b5Xp-ciZcQT1muYTf180SYUFHMVc/pub?gid=225335635&single=true&output=csv")
Sci10 <- read.csv("https://docs.google.com/spreadsheets/d/1HZJxmJ-n0dEdkNLlNox8lhUEkWdBigoycdB4Gh-mAlg/pub?gid=46470343&single=true&output=csv")

## Writting

Write14 <- read.csv("https://docs.google.com/spreadsheets/d/1rXrvkCAAFrO8_N-6YznQwuId0NTMOlo9OrI5wX18jJw/pub?gid=1285166754&single=true&output=csv")
Write13 <- read.csv("https://docs.google.com/spreadsheets/d/1EXWDmHLyXVjenck1AiXxxDRpqFEJEMuwt_XBLzTGozo/pub?gid=881482941&single=true&output=csv")
Write12 <- read.csv("https://docs.google.com/spreadsheets/d/1F4DNMNUAcV4U80NIJcLXzNYvZII8n2VdhnynId8WLpE/pub?gid=1358804477&single=true&output=csv")

#only 4-7
Write11 <- read.csv("https://docs.google.com/spreadsheets/d/1wuENaBv747xUCj5NGT9CtIM79QqMHzxK3NAWyqWmiEo/pub?gid=799752116&single=true&output=csv")

#only 4-7
Write10 <- read.csv("https://docs.google.com/spreadsheets/d/167c1jshh1156x-vJUE7L-R0EJTstuIegssfMwKwUEXk/pub?gid=156237836&single=true&output=csv")


```


## Eligible schools

Not all schools are eligible.  They must be served by an IOU.  Original file has odd formating.  So, saved and accessed via gdrive.

```{r}
#Non readable because of formating to make pretty.
#read.xlsx("http://www.oregon.gov/energy/SCHOOLS/Sb1149/documents/schools_list.xls")

Eligible <- read.csv("https://docs.google.com/spreadsheets/d/1WXNu3cNY4DyZ7yWXT2MPd3bd61M0Vmo2ty62Uux1TZU/pub?gid=1048793783&single=true&output=csv")
```

## Installed measures

```{r}
EEMs <-read.csv("https://docs.google.com/spreadsheets/d/15X3x4KCEx2-_JYLjq3gXhqS33EKBscW_LLPO-MggDFk/pub?gid=1734037341&single=true&output=csv")
```

## District demographics

Load dataframes for School performance from 2002-2016. The name for each dataframe is "Schoolxx" where xx is the last 2 digits of the ending academic year from which the data is for. The formating and the information provided in these school report cards slightly changes after 2012.   

```{r}
Schools16 <- read.csv("http://www.ode.state.or.us/data/reportcard/media/17/RCmediaSchoolsAggregate.csv")
Schools15 <- read.csv("http://www.ode.state.or.us/data/reportcard/media/16/RCmediaSchoolsAggregate.csv")
Schools14 <- read.csv("http://www.ode.state.or.us/data/reportcard/media/15/RCmediaSchoolsAggregate.csv")
Schools13 <- read.csv("http://www.ode.state.or.us/data/reportcard/media/14/RCmediaSchoolsAggregate.csv")
Schools12 <- read.csv("http://www.ode.state.or.us/data/reportcard/media/13/RCmediaSchoolsAggregate.csv")
Schools11 <- read.csv("http://www.ode.state.or.us/data/reportcard/media/12/RCmediaSchoolsAggregate.csv")
Schools10 <- read.csv("http://www.ode.state.or.us/data/reportcard/media/11/RCmediaSchoolsAggregate.csv")
Schools09 <- read.csv("http://www.ode.state.or.us/data/reportcard/media/10/RCmediaSchoolsAggregate.csv")
Schools08 <- read.csv("http://www.ode.state.or.us/data/reportcard/media/09/RCmediaSchools.csv")
Schools07 <- read.csv("http://www.ode.state.or.us/data/reportcard/media/08/RCmediaSchools.csv")
Schools06 <- read.csv("http://www.ode.state.or.us/data/reportcard/media/07/RCmediaSchools.csv")
Schools05 <- read.csv("http://www.ode.state.or.us/data/reportcard/media/06/RCmediaSchools.csv")
Schools04 <- read.csv("http://www.ode.state.or.us/data/reportcard/media/05/RCmediaSchools.csv")
Schools03 <- read.csv("http://www.ode.state.or.us/data/reportcard/media/04/RCmedia.csv")

```

# Linking Data

Lets get the specific data items we need for each district.  Please note that this is district level data so we need to get expenditures for all grade levels as well as construction expenditures.


```{r}
summary(DistrictFinance$FuncDesc)

```


The key fund descriptions are:


+ Plant Capital
    + Building Acquisition; Construction; and Improvement Services 
    + Facilities Acquisition and Construction 
    + Site Acquisition and Development Services

+ Operation and Maintenance of Plant Services 

+ Education Expenditures
    + English Second Language Programs 
    + High School Programs 
    + Intermediate Programs
    + Middle/Junior High Programs
    + Primary (K-3) 

Will go with the major category of instruction for Educational expenditures.

```{r}
DistrictFinance %>% 
  filter( FuncDesc == 'Building Acquisition; Construction; and Improvement Services' | FuncDesc == 'Facilities Acquisition and Construction' | FuncDesc == 'Site Acquisition and Development Services') %>% 
  mutate(ExpType = 'Capital') -> CapitalExpenditures

DistrictFinance %>% 
  filter( FuncDesc == 'Operation and Maintenance of Plant Services' ) %>%
  mutate( ExpType = 'Operations' ) -> PlantOperations

# Going for the larger Instructional Category
# 
# DistrictFinance %>% filter(FuncDesc == 'English Second Language Programs' | FuncDesc == 'High School Programs' |  FuncDesc == 'Intermediate Programs' |FuncDesc == 'Middle/Junior High Programs' | FuncDesc == 'Primary (K-3)') %>% mutate(ExpType = 'Education') -> EducationExpenditures

DistrictFinance %>% 
  filter( MajFuncDesc == 'Instruction' ) %>% 
  mutate( ExpType = 'Education' ) -> EducationExpenditures

KeyBudget <- bind_rows(list(CapitalExpenditures, PlantOperations, EducationExpenditures))


```


Turn this into a one row per district per year data frame

```{r}
library(tidyr)
KeyBudget %>% spread(FuncDesc, Amount) -> Budget

```




