---
title: "Test if Border District Identification Works"
output: html_notebook
---

Steps:

+ Get database of all schools in Oregon entered.
+ Purge schools that are Charter, immersion or magnet schools.
+ Find districts that have schools that are both eligible and not eligible, i.e., border districts.
+ Of that group see if there are schools with installed EE measures.
+ Find districts with treated schools
+ Find only schools in border districts that have treated schools

# All Schools

```{r}
# All schools with eligibility status
 
AllSchools <- read.csv("https://docs.google.com/spreadsheets/d/1WXNu3cNY4DyZ7yWXT2MPd3bd61M0Vmo2ty62Uux1TZU/pub?gid=1048793783&single=true&output=csv")

```

# Purging unusual schools

```{r}
library(dplyr)
library(stringr)


AllSchools %>% 
  mutate(Charter = str_detect(School.Name, "Charter")) %>%
  mutate(Immersion = str_detect(School.Name, "Immersion")) %>%
  mutate(Magnet = str_detect(School.Name, "Magnet")) -> AllSchools

summary(AllSchools)

```


Notice that school names are not unique.

Remove the charters and others.

```{r}
AllSchools %>% 
  filter (Charter == FALSE & Immersion == FALSE  & Magnet == FALSE) -> AllSchools
```

The notes indicate that some schools need to be purged

+ "Housed at"
+ "shared campus"
+ "not own"
+ "not owned"


```{r}
AllSchools %>%
  mutate(SpecialPurge = str_detect(Notes, "Housed at") |
           str_detect(Notes, "shared campus") |
           str_detect(Notes, "not own") | 
           str_detect(Notes, "not owned")) -> AllSchools

AllSchools %>% 
  filter (SpecialPurge == FALSE) -> AllSchools
```


# Find Border Districts

```{r}
AllSchools %>% 
  group_by(District.ID, District.Name) %>%
  count(SB.1149.Eligibility) %>%
  group_by(District.ID, District.Name) %>%
  count() %>%
  filter(nn >1) -> BorderDistricts

```

Only schools in Border Districts

```{r}
AllSchools %>%
  inner_join(BorderDistricts) -> BorderSchools

```

# Find the schools in this group with EEMs


All EEMs load
```{r}

EEMs <-read.csv("https://docs.google.com/spreadsheets/d/15X3x4KCEx2-_JYLjq3gXhqS33EKBscW_LLPO-MggDFk/pub?gid=1734037341&single=true&output=csv")


```


Find unique schools with EEMs

```{r}
EEMs %>% 
  select(FacilityODEId) %>%
  unique() -> TreatedSchoolIDs

TreatedSchoolIDs$Treated <- TRUE

```

Tag the border schools that have had EEMs

```{r}
left_join(BorderSchools, TreatedSchoolIDs, by = c("School.ID" = "FacilityODEId" )) %>%
  mutate(Treated = if_else(Treated == TRUE, TRUE, FALSE, missing=FALSE)) -> BorderSchools
  

```


# Districts with treated schools

```{r}
BorderSchools %>% 
  filter(Treated == TRUE) %>%
  group_by(District.ID, District.Name) %>%
  count(Treated) %>%
  select(District.ID,District.Name,n)
  arrange(n)

```

Keep only schools  within districts that have a school that has been treated.


```{r}
BorderSchools %>% 
  filter(Treated == TRUE) %>%
  group_by(District.ID, District.Name) %>%
  count(Treated) %>% 
  filter(n >0) %>%
  select(District.ID,District.Name)-> WithTreated

BorderSchools %>%
  semi_join(WithTreated) -> TreatedDistrictSchools

summary(TreatedDistrictSchools)
```
```{r}

table(TreatedDistrictSchools$SB.1149.Eligibility,TreatedDistrictSchools$Treated)

```

# Merge in treatment dates and details

# Merge in test scores

# Choose school characteristics for comparability


