---
title: "OregonSchools"
output: pdf_document
---

Load Libraries


```{r}
library(dplyr)
library(stringr)

```

Load dataframes for School performance from 2002-2016. The name for each dataframe is "Schoolxx" where xx is the last 2 digits of the ending academic year from which the data is for. The formating and the information provided in these school report cards slightly changes after 2012.   

```{r}
Schools16 <- read.csv("http://www.ode.state.or.us/data/reportcard/media/17/RCmediaSchoolsAggregate.csv")
Schools15 <- read.csv("http://www.ode.state.or.us/data/reportcard/media/16/RCmediaSchoolsAggregate.csv")
Schools14 <- read.csv("http://www.ode.state.or.us/data/reportcard/media/15/RCmediaSchoolsAggregate.csv")
Schools13 <- read.csv("http://www.ode.state.or.us/data/reportcard/media/14/RCmediaSchoolsAggregate.csv")
Schools12 <- read.csv("http://www.ode.state.or.us/data/reportcard/media/13/RCmediaSchoolsAggregate.csv")
Schools11 <- read.csv("http://www.ode.state.or.us/data/reportcard/media/12/RCmediaSchoolsAggregate.csv")
Schools10 <- read.csv("http://www.ode.state.or.us/data/reportcard/media/11/RCmediaSchoolsAggregate.csv")
Schools09 <- read.csv("http://www.ode.state.or.us/data/reportcard/media/10/RCmediaSchoolsAggregate.csv")
Schools08 <- read.csv("http://www.ode.state.or.us/data/reportcard/media/09/RCmediaSchools.csv")
Schools07 <- read.csv("http://www.ode.state.or.us/data/reportcard/media/08/RCmediaSchools.csv")
Schools06 <- read.csv("http://www.ode.state.or.us/data/reportcard/media/07/RCmediaSchools.csv")
Schools05 <- read.csv("http://www.ode.state.or.us/data/reportcard/media/06/RCmediaSchools.csv")
Schools04 <- read.csv("http://www.ode.state.or.us/data/reportcard/media/05/RCmediaSchools.csv")
Schools03 <- read.csv("http://www.ode.state.or.us/data/reportcard/media/04/RCmedia.csv")

```


Keep only relevant Columns for each dataframe such as school ID number, name, type, total enrollment, percent of regular attendance, and finally, student performance metrics. 

We're only looking at percent of students that pass math each academic year. Testing methods and available information for science and language arts/writing/reading are inconsistent over the years. Columns are named differently for dataframes from 2012 and earlier. 

These smaller dataframes are named "Pxx" where xx is the last 2 digits of the ending academic year from which the data is for.

```{r}

P16 <- Schools16 %>% subset(, c("School.ID","School.Name", "School.Type", "Total.Enrollment", "Pct.of.Students.Attending.Regularly", "Math.Pct.Met.2015.16"))

P15 <- Schools15 %>% subset(, c("School.ID", "Total.Enrollment", "Pct.of.Students.Attending.Regularly", "Math.Pct.Met.2014.15"))

P14 <- Schools14 %>% subset(, c("School.ID", "Total.Enrollment" , "Pct.of.Students.Attending.Regularly", "Math.Pct.Met.2013.14"))

P13 <- Schools13 %>% subset(, c("School.ID", "Total.Enrollment" , "Pct.of.Students.Attending.Regularly", "Math.Pct.Met.2012.13"))

P12 <- Schools12 %>% subset(, c("InstID", "DailyMmbrshpAvgAmt1", "AttdPct2", "AttdPct1","MathCompSchoolAllGradePct" ))

P11 <- Schools11 %>% subset(, c("InstID", "DailyMmbrshpAvgAmt1", "AttdPct2", "AttdPct1", "MathCompSchoolAllGradePct" ))

P10 <- Schools10 %>% subset(, c("InstID", "DailyMmbrshpAvgAmt1", "AttdPct2","AttdPct1", "MathCompSchoolAllGradePct"))

P09 <- Schools09 %>% subset(, c("InstID", "DailyMmbrshpAvgAmt1", "AttdPct2", "AttdPct1", "MathCompSchoolAllGradePct"))

P08 <- Schools08 %>% subset(, c("InstID", "DailyMmbrshpAvgAmt1", "AttdPct2", "AttdPct1", "MathCompSchoolAllGradePct" ))

P07 <- Schools07 %>% subset(, c("InstID", "DailyMmbrshpAvgAmt1", "AttdPct2", "AttdPct1", "MathCompSchoolAllGradePct"))

P06 <- Schools06 %>% subset(, c("InstID", "DailyMmbrshpAvgAmt1", "AttdPct2", "AttdPct1", "MathCompSchoolAllGradePct"))

P05 <- Schools05 %>% subset(, c("InstID", "DailyMmbrshpAvgAmt1", "AttdPct2", "AttdPct1", "MathCompSchoolAllGradePct"))

P04 <- Schools04 %>% subset(, c( "instid", "DailyMmbrshpAvgAmt1", "AttdPct2", "AttdPct1", "MathPSCompSchoolAllGradePct"))

P03 <- Schools03 %>% subset(,c("instid", "DailyMmbrshpAvgAmt1", "AttdPct4" , "AttdPct3", "AttdPct2" , "AttdPct1", "MathCompSchoolAllGradePct"))

```


Rename the columns in each dataframe so that they are consistent through all of them. The dataframe P03 does not have a SchoolType column (E,M,H).

```{r}
library(base)

names(P16) <- c("SchoolID", "SchoolName", "SchoolType", "Enrollment2016", "Attendance2016","Math2016")
P16$Math2016 <- as.numeric(as.character(P16$Math2016))

names(P15) <- c("SchoolID", "Enrollment2015", "Attendance15", "Math2015")
P15$Math2015 <- as.numeric(as.character(P15$Math2015))

names(P14) <- c("SchoolID", "Enrollment2014", "Attendance14", "Math2014")
P14$Math2014 <- as.numeric(as.character(P14$Math2014))

names(P13) <- c("SchoolID", "Enrollment2013", "Attendance13", "Math2013")
P13$Math2013 <- as.numeric(as.character(P13$Math2013))

names(P12) <- c("SchoolID", "Enrollment2012", "2Attendance12", "1Attendance12", "Math2012")
P12$Math2012 <- as.numeric(as.character(P12$Math2012))

names(P11) <- c("SchoolID", "Enrollment2011", "2Attendance11", "1Attendance11",  "Math2011")
P11$Math2011 <- as.numeric(as.character(P11$Math2011))

names(P10) <- c("SchoolID" , "Enrollment2010", "2Attendance10", "1Attendance10", "Math2010")
P10$Math2010 <- as.numeric(as.character(P10$Math2010))

names(P09) <- c("SchoolID", "Enrollment2009", "2Attendance09", "1Attendance09", "Math2009")
P09$Math2009 <- as.numeric(as.character(P09$Math2009))

names(P08) <- c("SchoolID", "Enrollment2008", "2Attendance08", "1Attendance08", "Math2008")
P09$Math2009 <- as.numeric(as.character(P09$Math2009))

names(P07) <- c("SchoolID", "Enrollment2007", "2Attendance07", "1Attendance07", "Math2007")
P07$Math2007 <- as.numeric(as.character(P07$Math2007))

names(P06) <- c("SchoolID", "Enrollment2006", "2Attendance06", "1Attendance06", "Math2006")
P06$Math2006 <- as.numeric(as.character(P06$Math2006))

names(P05) <- c("SchoolID", "Enrollment2005", "2Attendance05", "1Attendance05", "Math2005")
P05$Math2005 <- as.numeric(as.character(P05$Math2005))

names(P04) <- c("SchoolID", "Enrollment2004", "2Attendance05", "1Attendance05","Math2004")
P04$Math2004 <- as.numeric(as.character(P04$Math2004))

names(P03) <- c("SchoolID", "Enrollment2003", "4Attendance2003", "3Attendance2003", "2Attendance2003", "1Attendance2003","Math2003")
P03$Math2003 <- as.numeric(as.character(P03$Math2003))

```

Combine dataframes for each year into one single dataframe

```{r}

M16 <- merge(P16, P15, by = "SchoolID")
M15 <- merge(M16, P14, by = "SchoolID")
M14 <- merge(M15, P13, by = "SchoolID")
M13 <- merge(M14, P12, by = "SchoolID")
M12 <- merge(M13, P11, by = "SchoolID")
M11 <- merge(M12, P10, by = "SchoolID")
M10 <- merge(M11, P09, by = "SchoolID")
M09 <- merge(M10, P08, by = "SchoolID")
M08 <- merge(M09, P07, by = "SchoolID")
M07 <- merge(M08, P06, by = "SchoolID")
M06 <- merge(M07, P05, by = "SchoolID")
M05 <- merge(M06, P04, by = "SchoolID")
M04 <- merge(M05, P03, by = "SchoolID")



```

Filter out all high schools and only keep elementary and middle schools

```{r}

SchoolPerformance <- subset(M04, SchoolType %in% c("E","M"))

```


Create csv file for final combined dataframe


```{r}
write.csv(SchoolPerformance, file = "StudentPerformance.csv")

#eligible <- read.csv("http://www.oregon.gov/energy/SCHOOLS/Sb1149/documents/schools_list.xls")

```

