---
title: "The Effect of Energy Efficiency Measures on K12 Educational Performance"
output: beamer_presentation
bibliography: ../Bibliography.bib
citation_package: natbib
biblio-style: "apalike"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Abstract


This is early work.  Still much to do.

# Background

+ Cost effectiveness evaluation requires an evaluation of _all_ costs and _all_ benefits.
+ Not all jurisdictions use "participant benefits" in evaluation of cost effectiveness because they are not enjoyed by the rate payer.  
+ The "National Efficiency Screening Project" only gives guidance on including non-energy benefits as it explains the symmetry principal, "If you include participant benefits, include participant costs".
+ Some jurisdictions, e.g., The Energy Trust of Oregon, are reducing the technical complexity of evaluations to _reduce evaluation costs_

# "Reduced Technical Complexity"

+ Free ridership/drivership by survey, "Would you have done this without ETO incentives?"
    + Assume people tell the truth.
    + Economists don't believe that.
+ Simpler econometrics
    + No accounting for self-selection bias.
    + No accounting for sampling bias.
    + Stops  'futzing'
    
In short, the opposite of what economics is doing with program evaluation, but more inline with engineering approach.

# Why the Effect of Energy Efficency on Education is Interesting

+ Decisions to improve structures is a financial decision -- money saved on energy.
+ No financial gain, not allowed to participate.

Improvements in student/staff health and the lifetime effects of academic performance dwarf the value of energy savings.

# Daylighting (Sampling)

+ D. A. Kleiber and others. “Environmental Illumination and Human
Behavior: The Effects of Spectrum of Light Source on Human Performance in a
University Setting.” (1973)
    + 3 schools  
    + Movement to portables.
    + 17% drop

+ L. Heschong. “Daylighting in Schools: An Investigation into the
Relationship between Daylighting and Human Performance. Detailed Report.”
(1999).
    + "Data indicate students with the most classroom daylighting progressed 20 percent faster on math tests and 26 percent on reading tests in one year than those with the least."
    + No control for self-selection
    + ~20,000 schools

# 

+ L. Heschong, R. L. Wright and S. Okura. “Daylighting impacts on human
performance in school”. In: _Journal of the Illuminating Engineering
Society_ 31.2 (2002), pp. 101-114.
    + Positive effect. Scale is dubious.
    + Multiple school districts with various quality of daylighting.
    

+ M. H. Nicklas and G. B. Bailey. “Analysis of the Performance of
Students in Daylit Schools.” (1996).
    + "... [daylit] schools outperformed students attending artificially lighted
schools by 5 to 14 percent."

# Indoor Air Quality (Sampling)

+ M. J. Mendell and G. A. Heath. “Do indoor pollutants and thermal
conditions in schools influence student performance? A critical review of
the literature”. In: _Indoor air_ 15.1 (2005), pp. 27-52.
    + Not all that critical( Three studies of students, x2 from 70s)
    + Summary, something is there but we don't know what it is.

+ S. Moonie, D. A. Sterling, L. W. Figgs, et al. “The relationship
between school absence, academic performance, and asthma status”. In:
_Journal of School Health_ 78.3 (2008), pp. 140-148.
    + 3K students
    + More absences, lower test scores.
    + Asthma kids have more absences but not lower scores with same absences.

#

+ U. Haverinen-Shaughnessy, D. Moschandreas and R. Shaughnessy.
“Association between substandard classroom ventilation rates and students’
academic achievement”. In: _Indoor air_ 21.2 (2011), pp. 121-131.
    + Schools with less than 7.1 l/s/person, i.e., less than ASHRAE Standard 62 in 2004.
    + "...[1 l/s per person] increase in the ventilation rate within that range, the proportion of students passing standardized test (i.e., scoring satisfactory or above) is expected to increase by 2.9% (95%CI 0.9–4.8%) for math and 2.7% (0.5–4.9%) for reading."

# In Short

+ Students got stuck in the 60s-70s air conditioning daylighting vs air conditioning battle.
+ Air quality is important but we don't know exactly what part
    + Keep in mind that allergenic mold concentrations, colony forming bodies, can change by factor of 10 in a few hours.
    + Hard to measure.
      
# Why Oregon Schools?

+ Nice features that remove many, but not all, self-selection problems.
+ Decison making cutoffs are based on energy savings, not educational outcomes.

+ Still problems
    + Data shortcoming ... as we go along.
    + Standardized tests are not standard from year to year.
    + Building data is not so important to departments of education.

Now, on to the data ...

# SB 1149

+ [SB 1149 (1999)](https://www.oregonlegislature.gov/bills_laws/lawsstatutes/1999orLaw0865.html) was Oregon's deregulation bill.
    + It was part of the implementation of FERC 888.
    + Established a public purpose charge that funds thing like The Energy Trust of Oregon.
+ The 3% Public Purpose Charge was collected by _almost_ all the IOUs.
    + Pacificorp and PGE collect, but Idaho Power in Eastern Oregon does not.
    + Used for Energy Efficiency
    + 10% must be used in schools.
    
# Schools Program    

Schools using SB 1149 public purpose funds:

+ Complete energy audits of _buildings_, not districts, served by Pacificorp or PGE (with limitations);
+ Audits must be completed by an approved audit company;
+ Implement the approved Energy Efficiency Measures identified in the audits;
+ Report Energy Use Index data in the Schools Interactive Database each year; 


# Key Points

+ Not all districts are eligible.
+ Not all schools within districts are eligible.


Makes it easier to establish a natural control group for any treated school by using others in the district.

# Educational Funding in Oregon

Oregon Educational Funding has key provisions that makes using schools in other district more attractive than other states.

+ 1991 Oregon establishes an Equalization Formula:
    + The legislature establishes a biennial K12 State School Fund Budget
    + The State School Fund budget is distributed equally by student across the state, but adjusted for property tax collections.
    + $1 increase in property tax collected results in $1 less in State School Fund support.
    + Measure 5 and 50 property tax limitations put most of the power with the legislature.
    
+ Result, equal, per-student funding across the state with a few exceptions:
    + Short-term, 3 year, property tax operating levies.
    + Capital bonds
    + A few, depends on year, school districts with few students but lots of taxable property.

Much smaller inter-district funding differences than other states.

# School District Size

+ Oregon has 197 school districts for ~500K students.
    + Washington has 296 for ~1M students
+ The three largest school districts, Portland Public, Salem-Keizer, and  Beaverton are about 40K each.
    + Washington has 9 over 20K
+ The remainder are small.

Less opportunity for intra-district funding differences but need to watch the big three.

# Key Points

+ Management and salary levels may be different across districts 
+ There are fewer haves and have nots within and between districts than other states.

We can feel more comfortable using schools out of district for controls given equal funding.

# The Energy Efficiency Measures (EEMs)

Mostly what you would expect

+ Building Envelope, including windows and insulation.
+ HVAC Components  
+ HVAC Controls
+ Pumps, Motors and Drives
+ Domestic Hot Water
+ Lighting
+ Kitchen Equipment
+ Other (Pool Covers ...)

# What Should Strike you About the List

+ Most of the items you would never notice unless you were an expert looking  or listening for them.
+ There are a few that could have an impact on the educational environment.
    + Double or Triple Glazed Windows ( Sound, Moisture Control)
    + Cavity Insulation (Sound, Moisture Control)
    + Lighting Quality Improvements (Daylighting)


# Key Points

+ Treated schools may or may not have educationally impactful EEMs.
+ Treated schools could have a mix.

The non-impactful EEMs can provide a placebo effect robustness check on the effects of the impactful EEMs on educational performance.

# How to Measure Educational Impact

+ Some of the measures produce environmental improvements, less moisture, mold, we can look at the effect on attendance.
    + Recorded annually for each school.
    + Intermediate indicator.  The more often you go to school the better you do.
+ Environmental Improvements can reduce teacher absenteeism.
    + Counts not recorded but expenditures on substitutes are a budget item that can be observed.
+ Some produce better learning environments, less distraction.  We can use the annual standardized testing results.
    + Everyone in the same grade takes the same test, but the test, and the levels can be different from year-to-year.

# Key Points

+ Test scores are the main indicator but for some EEMs, absenteeism and attendance may be good predictors.
+ We have some candidates for instruments if we wish to take that route.


# Data Details

# Schools Program Audit Requirements

+ Non-educational buildings and those that are rented or will close in 5 years are ineligible.
+ Energy Audits are required
    + Whole Building audits, similar to ASHRAE Level 2.  Identify EEMs with 50 year payback or less.
    + Multi-component payback is allowed
    + Targeted Audits are acceptable for limited scope.
    + Target is 47/48 kBTU/SF/Year for elementary and 61/62 kBTU/SF/Year for High Schools 
+ Multi-competent payback calculations are allowed.

# School Program Implementation

+ The maximum amount of PPC funds reimbursed will be capped at the total annual savings multiplied by the Measure Life caped at cost.
+ Common for some cost to not be funded.
+ Commissioning is required for:
    + All boiler or chiller measures exceeding $100,000
    + All other HVAC measures and all HVAC controls measures exceeding $50,000
    + All lighting control measures exceeding $100,000
    + Other measures in which commissioning is critical for successful implementation and operation of the measure, as deemed appropriate by the auditor.

# Required Annual Reporting

+ Annual energy expenses by fuel type
+ Square footage
+ Hours of operation

# Summary of Measures

```{r InstallByYear, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
EEMs <-read.csv("https://docs.google.com/spreadsheets/d/15X3x4KCEx2-_JYLjq3gXhqS33EKBscW_LLPO-MggDFk/pub?gid=1734037341&single=true&output=csv")

library(dplyr)
library(knitr)

EEMs %>% select(Year  = ImplementedYear) %>% 
  group_by(Year) %>%
   summarize(Installations = n()) %>%
  kable()
```

Note spikes in installation.

# Types of Installed Measures

```{r InstallTypes, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
EEMs %>% select(EEM = MeasureClassDetailDescription) %>%
  group_by(EEM) %>%
  summarize(Installations = n()) %>% 
  kable()

```

# Comments

+ Controls are frequently occupancy sensors and day-lighting controls
+ Fixtures are described in detail later as gym, exterior, etc.

# Electric Utilities

![BPA 2015](https://www.bpa.gov/news/pubs/maps/OregonUtils.pdf)

# Oregon is a mix

+ Most of population is served by IOUs
+ Large tracts of COUs
+ Only Pacificorp and PGE schools are eligible.

# Districts with Eligible Schools

```{r BuildBorderDistricts, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

AllSchools <- read.csv("https://docs.google.com/spreadsheets/d/1WXNu3cNY4DyZ7yWXT2MPd3bd61M0Vmo2ty62Uux1TZU/pub?gid=1048793783&single=true&output=csv")
library(dplyr)
library(stringr)


AllSchools %>% 
  mutate(Charter = str_detect(School.Name, "Charter")) %>%
  mutate(Immersion = str_detect(School.Name, "Immersion")) %>%
  mutate(Magnet = str_detect(School.Name, "Magnet")) -> AllSchools

AllSchools %>% 
  filter (Charter == FALSE & Immersion == FALSE  & Magnet == FALSE) -> AllSchools

AllSchools %>%
  mutate(SpecialPurge = str_detect(Notes, "Housed at") |
           str_detect(Notes, "shared campus") |
           str_detect(Notes, "not own") | 
           str_detect(Notes, "not owned")) -> AllSchools

AllSchools %>% 
  filter (SpecialPurge == FALSE) -> AllSchools

AllSchools %>% 
  group_by(District.ID, District.Name) %>%
  count(SB.1149.Eligibility) %>%
  group_by(District.ID, District.Name) %>%
  count() %>%
  filter(nn >1) -> BorderDistricts
```

```{r MapBoarderDistricts, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

library(maptools)
Sto <- gpclibPermit()

Schools <- readShapePoly("../MapData/cb_2016_41_unsd_500k/cb_2016_41_unsd_500k.shp")

library(ggmap)

mapImage <- get_map(location = "oregon", zoom = 7, maptype = "toner")


library(RColorBrewer)
colors <- brewer.pal(9, "BuGn")

# Subset SchoolBound so only eligible districts
# 


Split <- function(x) { last(str_split(x, boundary("word"))[[1]])
}

SplitV <- Vectorize(Split)



DistNumbers <- SplitV(as.character(BorderDistricts$District.Name))

# Fix so that only the eligible districts show.
#Schools@data$NAME

SchoolBound <- fortify(Schools)


ggmap(mapImage) 
# +
#   geom_polygon(aes(x = long,
#                    y = lat,
#                    group = group),
#                data = SchoolBound,
#                color = colors[9],
#                fill = colors[6],
#                alpha = 0.5) +
#   labs(x = "Longitude",
#        y = "Latitude")
# 

```

 
# How Many District? Schools?

```{r EligibleSchools, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

AllSchools <- read.csv("https://docs.google.com/spreadsheets/d/1WXNu3cNY4DyZ7yWXT2MPd3bd61M0Vmo2ty62Uux1TZU/pub?gid=1048793783&single=true&output=csv")

AllSchools %>% 
  mutate(Charter = str_detect(School.Name, "Charter")) %>%
  mutate(Immersion = str_detect(School.Name, "Immersion")) %>%
  mutate(Magnet = str_detect(School.Name, "Magnet")) -> AllSchools

AllSchools %>% 
  filter (Charter == FALSE & Immersion == FALSE  & Magnet == FALSE) -> AllSchools

AllSchools %>%
  mutate(SpecialPurge = str_detect(Notes, "Housed at") |
           str_detect(Notes, "shared campus") |
           str_detect(Notes, "not own") | 
           str_detect(Notes, "not owned")) -> AllSchools

AllSchools %>% 
  filter (SpecialPurge == FALSE) -> AllSchools

AllSchools %>% 
  group_by(District.ID, District.Name) %>%
  count(SB.1149.Eligibility) %>%
  group_by(District.ID, District.Name) %>%
  count() %>%
  filter(nn >1) -> BorderDistricts

AllSchools %>%
  inner_join(BorderDistricts) -> BorderSchools

EEMs %>% 
  select(FacilityODEId) %>%
  unique() -> TreatedSchoolIDs

TreatedSchoolIDs$Treated <- TRUE

left_join(BorderSchools, TreatedSchoolIDs, by = c("School.ID" = "FacilityODEId" )) %>%
  mutate(Treated = if_else(Treated == TRUE, TRUE, FALSE, missing=FALSE)) -> BorderSchools
  
BorderSchools %>% 
  filter(Treated == TRUE) %>%
  group_by(District.ID, District.Name) %>%
  count(Treated) %>% 
  filter(n >0) %>%
  select(District.ID,District.Name)-> WithTreated

BorderSchools %>%
  semi_join(WithTreated) -> TreatedDistrictSchools

##
AllSchools %>% 
  filter(SB.1149.Eligibility == "YES") %>%
  select(District.ID) %>%
  unique() %>%
  nrow() -> EligibleDistricts

AllSchools %>% 
  filter(SB.1149.Eligibility == "YES") %>%
  select(School.ID) %>%
  unique() %>%
  nrow() -> EligibleSchools
```

+ Eligible 
    + Districts: `r EligibleDistricts` 
    + Schools: `r EligibleSchools`
+ Border (Eligible and ineligible schools in district)
    + Districts: `r nrow(BorderDistricts)`
    + With treated schools: `r nrow(WithTreated)`
+ Schools in Border Districts : `r nrow(BorderSchools)`
    + Eligible untreated: `r  table(BorderSchools$SB.1149.Eligibility)[2] - table(BorderSchools$Treated)[2]`
    + Eligible treated: `r table(BorderSchools$Treated)[2]`
    + Ineligible: `r table(BorderSchools$SB.1149.Eligibility)[1]`
+ Schools in Border Districts with treated schools : `r nrow(TreatedDistrictSchools)`
    + Eligible untreated: `r  table(TreatedDistrictSchools$SB.1149.Eligibility)[2] - table(TreatedDistrictSchools$Treated)[2]`
    + Eligible treated: `r table(TreatedDistrictSchools$Treated)[2]`
    + Ineligible: `r table(TreatedDistrictSchools$SB.1149.Eligibility)[1]`    
    


# Getting at Causality, the Identification Problem

# Methods

+ Corsened exact matching
    + Define what you mean by close enough match
    + Find the close enough match on observable dimensions
    + Excluded treated with no match
+ Synthetic control method
    + Mix of differences in differences and matching


# Matching Options

Control group composed of ineligible schools

+ Only within district or not
+ Variables to use with matching
+ Additional variable for matches outside of district


# How Different are Treated, Untreated and Ineligible schools?

Ideally, there should be no _obserable_ difference. As close as you can get to random assignment.

+ Dimensions:
    + Student population, free and reduced rate, ethnicity, etc.
    + Test scores
    + Facility characteristics
+ Problems:
    + Oregon Department of Education discontinued the School Facilities Report in 2002. Only have built and remodel dates before 2001.
    + Material that was in the report is only available for audited buildings in schools.
    + Schools are collections of buildings, not one, very common to have multiple buildings with a variety of vintages. 
    + Elementary near my house has 4.



